### 2021-10-25 paper/pldi22/formal.tex
More details in formalization

    \text{Primive Values} & \pval &\in& \pvalset &=&
    \boolset \uplus \strset \uplus \cdots\\

    \text{JavaScript ASTs} & \tree &\in& \treeset &::=& \\
  \end{array}
\]

States $\stset$ consist of labels $\labset$, calling context stacks
$\ctxtset^*$, heaps $\heapset$, and environments $\envset$.  A calling context
$\ctxt \in \ctxtset$ consists of a label denoting the return point, a caller's
environment, and a return variable.  A heap $\heap \in \heapset$ is a finite
mapping from addresses to objects, and an object $\obj \in \objset$ is a finite
mappings from strings to values.  Each object allocation $\kwcl \kwcr$ creates a
unique address $\addr \in \addrset$ different from existing addresses.  An
environment $\env \in \envset$ is a finite mapping from variables to values. A
value $\val \in \valset$ is a primitive value $\pval \in \pvalset$ (e.g., a
boolean value $\bool \in \boolset$ or a string value $\str \in \strset$), an
address $\addr \in \addrset$, a JavaScript AST $\tree \in \treeset$, or a
function $f \in \funcset$.

@@

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog = (\istset,
\getinst, \getnext)$ is defined as follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \ist \in \istset \wedge \ist \trans^* \st \}
\]
where $\trans^*$ denotes one or more repetition of $\trans$, and $\st \trans
\st'$ if and only if $\st = (\lab, \_, \_, \_)$ and $\sem{\getinst(\lab)}(\st) =
\st'$. Now, we define the denotational semantics of instructions $\sem{\inst}:
\stset \rightarrow \stset$ and expressions $\sem{\expr}: \stset \rightarrow
\stset \times \valset$.

\todo function semantics

\paragraph{Instructions}
\[
  \framebox{$\sem{\inst}: \stset \rightarrow \stset$}
\]
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \sem{\varx = \expr}(\st) =
      (\getnext(\lab), \ctxts, \heap, \env[\varx \mapsto \val])
    \]
    where
    \[
      \sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)
    \]

@@

  \item \underline{Branches}:
    \[
      \sem{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\st) =
      \left\{
        \begin{array}{ll}
          (\lab_\vart, \ctxts, \heap, \env) & \text{if} \; \val = \true\\
          (\lab_\varf, \ctxts, \heap, \env) & \text{if} \; \val = \false\\
        \end{array}
      \right.
    \]
    where
    \[
      \sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)
    \]

  \item \underline{Returns}:
    \[
      \sem{\kwret \; \expr}(\st) = (\lab, \ctxts, \heap, \env[\varx \mapsto
      \val])
    \]
    where
    \[
      \sem{\expr}(\st) = ((\_, (\lab, \env, \varx) :: \ctxts, \heap, \_), \val)
    \]
\end{itemize}

\paragraph{Expressions}
\[
  \framebox{$\sem{\expr}: \stset \rightarrow \stset \times \valset$}
\]
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \sem{\pval}(\st) =
      (\st, \pval)
    \]

  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\st) =
      (\st_n, \op(\val_1, \cdots, \val_n))
    \]
    where
    \[
      \st_0 = \st \wedge
      \forall 1 \leq k \leq n. \; \sem{\expr_k}(\st_{k-1}) = (\st_k, \val_k)
    \]

  \item \underline{Object Allocations}:
    \[
      \sem{\kwcl \kwcr}(\st) =
      ((\lab, \ctxts, \heap[\addr \mapsto \varnothing], \env), \addr)
    \]
    where
    \[
      \st = (\lab, \ctxts, \heap, \env) \wedge
      \addr \not\in \text{Domain}(\heap)
    \]

  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) =
      (\st, \env(\varx))
    \]
    where
    \[
      \st = (\_, \_, \_, \env)
    \]

  \item \underline{Field Lookups}:
    \[
      \sem{\expr_0 \kwsl \expr_1 \kwsr}(\st) =
      (\st'', \val')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr_0}(\st) &=& (\st', \addr) &\wedge\\
        \sem{\expr_1}(\st') &=& (\st'', \str) &\wedge\\
        \st'' &=& (\_, \_, \heap, \_) &\wedge\\
        \obj &=& \heap(\addr) &\wedge\\
        \val' &=& \obj(\str)\\
      \end{array}
    \]
\end{itemize}

>>>>

    \text{Primive Values} & \pval &\in& \pvalset &=&
    \boolset \uplus \intset \uplus \strset \uplus \cdots\\

    \text{JavaScript ASTs} & \tree &\in& \treeset\\
  \end{array}
\]

States $\stset$ consist of labels $\labset$, calling context stacks
$\ctxtset^*$, heaps $\heapset$, and environments $\envset$.  A calling context
$\ctxt \in \ctxtset$ consists of a label denoting the return point, a caller's
environment, and a return variable.  A heap $\heap \in \heapset$ is a finite
mapping from addresses to objects, and an object $\obj \in \objset$ is a finite
mappings from strings to values.  Each object allocation $\kwcl \kwcr$ creates a
unique address $\addr \in \addrset$ different from existing addresses.  An
environment $\env \in \envset$ is a finite mapping from variables to values. A
value $\val \in \valset$ is a primitive value $\pval \in \pvalset$ (e.g., a
boolean value $\bool \in \boolset$, an integer $k \in \intset$, or a string
$\str \in \strset$), an address $\addr \in \addrset$, a JavaScript AST $\tree
\in \treeset$, or a function $f \in \funcset$.

@@

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog = (\istset,
\getinst, \getnext)$ is defined as follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \ist \in \istset \wedge \ist \trans^* \st \}
\]
where $\trans^*$ denotes one or more repetition of $\trans$, and $\st \trans
\st'$ if and only if $\st = (\lab, \_, \_, \_)$ and $\sem{\getinst(\lab)}(\st) =
\st'$. Now, we define the denotational semantics of expressions, instructions,
and functions.

\paragraph{Expressions} We define the denotational semantics of expressions with
the following form:
\[
  \framebox{$\sem{\expr}: \stset \rightarrow \stset \times \valset$}
\]
For each expression $\expr \in \exprset$, its semantics $\sem{\expr}$ takes a
state and returns a pair of an updated state and a value as the result of
expression. We define five different cases in the semantics of expressions as
follows:
\begin{itemize}
  \item \underline{Primitive Values}:
    \[
      \sem{\pval}(\st) =
      (\st, \pval)
    \]

  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\st) =
      (\st_n, \op(\val_1, \cdots, \val_n))
    \]
    where $\st_0 = \st \wedge \forall 1 \leq k \leq n. \;
    \sem{\expr_k}(\st_{k-1}) = (\st_k, \val_k)$

  \item \underline{Object Allocations}:
    \[
      \sem{\kwcl \kwcr}(\st) =
      ((\lab, \ctxts, \heap[\addr \mapsto \varnothing], \env), \addr)
    \]
    where $\st = (\lab, \ctxts, \heap, \env) \wedge \addr \not\in
    \text{Domain}(\heap)$

  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) =
      (\st, \env(\varx))
    \]
    where $\st = (\_, \_, \_, \env)$

  \item \underline{Field Lookups}:
    \[
      \sem{\expr_0 \kwsl \expr_1 \kwsr}(\st) =
      (\st'', \obj(\str))
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr_0}(\st) &=& (\st', \addr) &\wedge\\
        \sem{\expr_1}(\st') &=& (\st'', \str) &\wedge\\
        \st'' &=& (\_, \_, \heap, \_) &\wedge\\
        \obj &=& \heap(\addr)\\
      \end{array}
    \]
\end{itemize}

\paragraph{Instructions} We define the denotational semantics of instructions
with the following form:
\[
  \framebox{$\sem{\inst}: \stset \rightarrow \stset$}
\]
For each instruction $\inst \in \instset$, its semantics $\sem{\inst}$ takes a
state and returns an updated state. We define five different cases in the
semantics of instructions as follows:

\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \sem{\varx = \expr}(\st) =
      (\getnext(\lab), \ctxts, \heap, \env[\varx \mapsto \val])
    \]
    where $\sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)$

@@

  \item \underline{Branches}:
    \[
      \sem{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\st) =
      \left\{
        \begin{array}{ll}
          (\lab_\vart, \ctxts, \heap, \env) & \text{if} \; \val = \true\\
          (\lab_\varf, \ctxts, \heap, \env) & \text{if} \; \val = \false\\
        \end{array}
      \right.
    \]
    where $\sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)$

  \item \underline{Returns}:
    \[
      \sem{\kwret \; \expr}(\st) = (\lab, \ctxts, \heap, \env[\varx \mapsto
      \val])
    \]
    where $\sem{\expr}(\st) = ((\_, (\lab, \env, \varx) :: \ctxts, \heap, \_),
    \val)$
\end{itemize}

\paragraph{Functions} Moreover, we define the denotational semantics of
functions with the following form:
\[
  \framebox{$\sem{\func}: \heapset \times \valset^* \rightarrow \heapset \times
  \valset$}
\]
For each function $\func \in \funcset$, its semantics $\sem{\func}$ takes a heap
and values as arguments, and it returns a pair of an updated heap and a return
value as follows:
\[
  \sem{\kwdef \; \varf \; \kwrl \varp_1, \cdots, \varp_n \kwrr \lab_\varf}
  (\heap, (\val_1, \cdots, \val_n)) = (\heap'', \val'')
\]
where
\[
  \begin{array}{l@{~}c@{~}ll}
    \st &=& (\lab_\varf, \epsilon, \heap, \env) &\wedge\\
    \env &=& [\varp_1 \mapsto \val_1, \cdots, \varp_n \mapsto \val_n] &\wedge\\
    \st \trans^* \st' &=& (\lab', \epsilon, \heap', \_) &\wedge\\
    \getinst(\lab') &=& \kwret \; \expr' &\wedge\\
    \sem{\expr'}(\st') &=& (\st'', \val'') &\wedge\\
    \st'' &=& (\_, \epsilon, \heap'', \_)\\
  \end{array}
\]
